

### Step 1: Load Required Libraries

```{r}
# Install required packages
#install.packages("e1071")     # For SVM
#install.packages("caret")     # For model evaluation
#install.packages("ggplot2")   # For visualization (optional)
```

Load the libraries:
```{r}
library(e1071)
library(caret)
library(ggplot2)
```

### Step 2: Load and Explore the Dataset
The Iris dataset is built into R, so you can load it directly:
```{r}
# Load dataset
data(iris)

# Inspect the data
head(iris)
summary(iris)
str(iris)
```

### Step 3: Data Preprocessing
Split the data into training and testing sets:
```{r}
# Set seed for reproducibility
set.seed(123)

# Split the data (70% training, 30% testing)
trainIndex <- createDataPartition(iris$Species, p = 0.7, list = FALSE)
trainData <- iris[trainIndex, ]
testData <- iris[-trainIndex, ]
```

### Step 4: Train an SVM Model
Train an SVM model on the training dataset:
```{r}
# Train the SVM model
svm_model <- svm(Species ~ ., data = trainData, kernel = "radial")

# View the model summary
summary(svm_model)
```

### Step 5: Make Predictions
Use the trained model to predict on the test dataset:
```{r}
# Make predictions
predictions <- predict(svm_model, testData)

# View predictions
table(predictions, testData$Species)
```

### Step 6: Evaluate the Model
Calculate the accuracy and visualize the results:
```{r}
# Confusion matrix
confusionMatrix(predictions, testData$Species)

# Plot SVM results (optional visualization for 2D projection)
plot(svm_model, iris, Sepal.Length ~ Sepal.Width)
```

### Step 7: Hyperparameter Tuning (Optional)
Use cross-validation to find the best hyperparameters for the SVM model:
```{r}
# Tune SVM parameters
tune_result <- tune.svm(Species ~ ., data = trainData,
                        kernel = "radial",
                        cost = 2^(-1:2),
                        gamma = 2^(-1:1))

# View the best model
summary(tune_result)

# Retrain the best model
best_model <- tune_result$best.model
```

### Step 8: Evaluate the Tuned Model
Test the performance of the tuned model:
```{r}
# Predictions using the best model
tuned_predictions <- predict(best_model, testData)

# Confusion matrix
confusionMatrix(tuned_predictions, testData$Species)
```

### Step 9: Visualize Results (Optional)
Visualize the decision boundaries for better understanding:
```{r}
# Visualization with ggplot2
ggplot(data = iris, aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
  geom_point(size = 3) +
  theme_minimal()
```

